---
itle: "Reverse Engineering Project"
author: "Stella S. Canino, Wade DeVinney, and Miles Grovic"
date: "October 20, 2024"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r echo=FALSE, message=FALSE}

```

```{r}
# Turn off scientific notation
options(scipen=999)

library(tidyverse)
library(janitor)
library(lubridate)

```

```{r}
# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

april_june24 <- read.csv("data/APRIL-JUNE-2024-SOD-DETAIL-GRID-FINAL.csv")

jan_march24 <- read.csv("data/JAN-MAR-2024-SOD-DETAIL-GRID-FINAL.csv")

# Clean required data and prepare for analysis if needed. 
april_june24$BUDGET.OBJECT.CODE <- as.character(april_june24$BUDGET.OBJECT.CODE)
jan_march24$BUDGET.OBJECT.CODE <- as.character(jan_march24$BUDGET.OBJECT.CODE)

# Now bind the rows together
us_rep_expense24 <- bind_rows(april_june24, jan_march24)

glimpse(us_rep_expense24)

```

```{r}
us_rep_expense24 <- us_rep_expense24 |>
  mutate(
    PERFORM.START.DT = as.Date(PERFORM.START.DT, format = "%d-%b-%y"),
    PERFORM.END.DT = as.Date(PERFORM.END.DT, format = "%d-%b-%y"),
    TRANSACTION.DATE = as.Date(TRANSACTION.DATE, format = "%d-%b-%y")
  )

glimpse(us_rep_expense24)

```


Shows basic exploratory analysis to demonstrate an understanding of the dataset,
including the number of rows and columns, any obvious limitations or flaws and
any reasons why it might not be able to answer the questions you've posed last week.

Describe the limits of the data in terms of timeframe, what is and is not included
and any codes or other information you'd need to find definitions for, and what 
you've done to better define it.

## Our combined dataset is 18 columns with 283,590 rows. The limits on timeframe
is that there's an inconsistency in the start data and perform date of the
expenditure. This means that we don't have a specific hold on the exact date
that the congressmen incurred that expense. Furthermore, the dataset column of
transaction date has a lot of NAs, which can cause us to not know the specifics
of these expenses so know whether or not congressmen used it during their primary
runs. 

One: There may be issues answering this question with our dataset since not all 
transactions have dates attached to them. This might make estimating the amount of
spending each program is making inaccurate.

Two: For our question about top spenders on travel we will have to focus on either
ground transportation or land transportation due to the fact that there are various 
descriptions of travel. For example, “TRANSPORATION OF THINGS TOTALS” could be anything,
so we need to be specific on one type of transportation

Three: For our question on who spent the most on advertising the data is clean
and specific that we can get a calculation on it. 

Four: For question four, the data is clean enough to see who had the most 
expenditures within the dataframe. Because we joined the two dataframes together, we can easily keep a running tally of who has spent what money thus far. 
The way we can express it in this data is with the sum of “AMOUNT” grouped by office.

Five: There wouldn’t be any issues for this question with the dataset. 
All of the descriptions are listed and they are well documented.

