---
itle: "Data Analysis Project"
author: "Stella S. Canino, Wade DeVinney, and Bao Bao"
date: "October 20, 2024"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
# Turn off scientific notation
options(scipen=999)

library(tidyverse)
library(janitor)
library(lubridate)

```

```{r}
# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

april_june24 <- read.csv("data/APRIL-JUNE-2024-SOD-DETAIL-GRID-FINAL.csv")

jan_march24 <- read.csv("data/JAN-MAR-2024-SOD-DETAIL-GRID-FINAL.csv")

# Clean required data and prepare for analysis if needed. 
april_june24$BUDGET.OBJECT.CODE <- as.character(april_june24$BUDGET.OBJECT.CODE)
jan_march24$BUDGET.OBJECT.CODE <- as.character(jan_march24$BUDGET.OBJECT.CODE)

# Now bind the rows together
us_rep_expense24 <- bind_rows(april_june24, jan_march24)

glimpse(us_rep_expense24)

```

```{r}
us_rep_expense24 <- us_rep_expense24 |>
  mutate(
    PERFORM.START.DT = as.Date(PERFORM.START.DT, format = "%d-%b-%y"),
    PERFORM.END.DT = as.Date(PERFORM.END.DT, format = "%d-%b-%y"),
    TRANSACTION.DATE = as.Date(TRANSACTION.DATE, format = "%d-%b-%y")
  )

glimpse(us_rep_expense24)

```


Shows basic exploratory analysis to demonstrate an understanding of the dataset,
including the number of rows and columns, any obvious limitations or flaws and
any reasons why it might not be able to answer the questions you've posed last week.

Describe the limits of the data in terms of timeframe, what is and is not included
and any codes or other information you'd need to find definitions for, and what 
you've done to better define it.

## Our combined dataset is 18 columns with 283,590 rows. The limits on timeframe
is that there's an inconsistency in the start data and perform date of the
expenditure. This means that we don't have a specific hold on the exact date
that the congressmen incurred that expense. Furthermore, the dataset column of
transaction date has a lot of NAs, which can cause us to not know the specifics
of these expenses so know whether or not congressmen used it during their primary
runs. 

One: There may be issues answering this question with our dataset since not all 
transactions have dates attached to them. This might make estimating the amount of
spending each program is making inaccurate.

Two: For our question about top spenders on travel we will have to focus on either
ground transportation or land transportation due to the fact that there are various 
descriptions of travel. For example, “TRANSPORATION OF THINGS TOTALS” could be anything,
so we need to be specific on one type of transportation

Three: For our question on who spent the most on advertising the data is clean
and specific that we can get a calculation on it. 

Four: For question four, the data is clean enough to see who had the most 
expenditures within the dataframe. Because we joined the two dataframes together, we can easily keep a running tally of who has spent what money thus far. 
The way we can express it in this data is with the sum of “AMOUNT” grouped by office.

Five: There wouldn’t be any issues for this question with the dataset. 
All of the descriptions are listed and they are well documented.

**DATA ANALYSIS 4**

In this fourth deliverable, we decided clean our answers for Q2, Q3, Q4 and Q5, 


```{r}
Q1 # Who increased their expenditures the most in the months leading up to the primary election?

#- need to filter data. also what type of expenditure are we looking for??
# - need to clarify since this is a lot of information.
#- this will be done with my group

US_congresspeeps_24 |>
  group_by() |>
  summarise()

```




Q2 Who is the top spender for traveling (airplanes, cars, etc)  in what is of  2024?

The top spender for what is 2024 is HON. LANCE GOODEN of Texas, who started 
his role in 2023 with $117,593.57. 

As this is an election year, such high spending is noteworthy, raising questions
about the use of public funds during a time of heightened political scrutiny.

```{r}
#Q2 Who is the top spender for traveling (airplanes, cars, etc)  in what is of  2024?

filtered_data <- US_congresspeeps_24 |> 
  filter(
    SORT.SUBTOTAL.DESCRIPTION == "TRAVEL") 
  
filtered_data <- filtered_data |> 
  filter(!DESCRIPTION %in% c("MEALS", "LODGING", "TRAVEL TOTALS:", "WI-FI ON TRAVEL"))

filtered_data |>
  group_by(ORGANIZATION) |>
 summarise(Total_Amount = sum(AMOUNT, na.rm = TRUE)) |> 
  arrange(desc(Total_Amount))   


```

Q3: Who is spending the most money on advertising?

The person who spent the most on advertisements was Shri Thanedar of Michigan with
$584,942.46. Next is Claudia Tenney of New York, and Carlos Gimenez of Florida.
An important thing to note is that in Q5 about the companies/people who were 
the vendors of the transactions SHRI THANEDAR	appeared with $344,313. 

#This is questionable and worth investigating why this congressmen is also a vendor
that's being given money as well.

```{r}
#Q3 Who is spending the most money on advertising?
 
US_congresspeeps_24 |>
  filter(DESCRIPTION == "ADVERTISEMENTS") |> 
  group_by(ORGANIZATION) |>                
  summarise(Total_Amount = sum(AMOUNT, na.rm = TRUE)) |> 
  arrange(desc(Total_Amount))   

```


Question 4:  Who is the top spender based on the data available for 2024, and what have they spent the most on?

#We filtered using string detect by legislative year, and "HON" to get the
correct time frame and so it only shows the congressmen. This is part 4A.

#For 4B, I filtered out using string detect the answers that were
totals of the expenses, transfers, and outside as a way to clean up the answers.
#We did this as a result of the blank answers from Q5 to make this much easier
to read the data. We then grouped by and summarized by the Description of his reimbursements. 

#I think the fact that he spent so much money on advertisements this year, which
is an election year is interesting. Mooney is from West Virginia, and he was
competing for the Senate seat of Manchin. He lost the primary to Justice. 

#It would be an interesting story idea to compare Mooney's advertisement expenses
before election year and now are different, and what are the ads he's pushing out.
Since it's illegal to use this money to help reelection campaings the publicity
of publishing what Mooney did in his term as representative could be a tactic for
his primary race.

A. 2024 HON. ALEXANDER X. MOONEY $510,7457.92. 
B.Mooney spent $276,950.18 on advertisements, which is his largest expense. 

```{r}
#4A
US_congresspeeps_24 <- us_rep_expense24 |> filter(FISCAL.YEAR.OR.LEGISLATIVE.YEAR == "LY2024") |> 
    filter(str_detect(ORGANIZATION, "HON"))

US_congresspeeps_24 |> 
  group_by(ORGANIZATION) |>
  summarise(Total_Amount = sum(AMOUNT, na.rm = TRUE)) |>
  arrange(desc(Total_Amount))
```

```{r}
#4B
Mooney24 <- us_rep_expense24 |>
  filter(ORGANIZATION == "2024 HON. ALEXANDER X. MOONEY")

Mooney24 |> 
  group_by(DESCRIPTION) |>
summarise(Total_Amount = sum(AMOUNT, na.rm = TRUE)) |>
  arrange(desc(Total_Amount)) |> filter(!str_detect(DESCRIPTION,"TOTALS")) |>
  filter(!str_detect(DESCRIPTION,"(TRANSFER)")) |>
   filter(!str_detect(DESCRIPTION,"(OUTSIDE)"))



```

## Question 5: Who is the top recipient of government money?

##The top recepient is the UNITED STATES POSTAL SERVICE, CITIBANK, and
AMPLIFY INC. It's important to know that we have $1,061,174,000 that has
no identifiable vendor, which is a worrying.

##UPDATE: We were able to identify that the majority of the NAs were the total
sum of the different values. 

#Now that we created a new dataframe with all the identifiable vendors, we
still see that the top recepeint is the U.S. Postal Service(1) and Citibank(2),LEIDOS DIGITAL SOLUTIONS INC(4) which isn't very interesting.

#However, the interesting top vendors are Amplify Inc(3) and AEJ Group LLC (5), and
The Franking Group (7). These three companies are related to media relations and
advertising firms that are working with congressmen.

#This is newsworthy since we see this pattern from Q4 and Q5 that advertising-related
expenses are the most that are being reimbursed in this legislative year.


```{r}
valid_vendors |> 
  group_by(VENDOR.NAME) |>
  summarise(Total_Amount = sum(AMOUNT, na.rm = TRUE)) |>
    arrange(desc(Total_Amount))
```

**Code for Q5 to figure out what's wrong**

```{r}
#This is our way to identify what's going in one with the NA in Q5. 

# Separate rows with missing or blank vendor names
missing_vendors <- US_congresspeeps_24 |> 
  filter(is.na(VENDOR.NAME) | str_trim(VENDOR.NAME) == "")

# Examine rows with valid vendor names
valid_vendors <- US_congresspeeps_24 |> 
  filter(!is.na(VENDOR.NAME) & str_trim(VENDOR.NAME) != "")

#There are more NA vendors that named vendors. --> with valid vendors having 136,086 obs.
```

```{r}

missing_vendors |>
  group_by(DESCRIPTION) |>
  summarise(Total_Amount = sum(AMOUNT, na.rm = TRUE)) |>
  arrange(desc(Total_Amount))

#OK! I(Stella) figured it out. The majority of the NAs are the totals
#of the reimbursements that make up the entirety of page 1 of this. 

#The other big group that gives us NAs are "transfers,"which I have no idea.

#Maintenance/Repairs is another group which means that it was probably internal reimbursement within the House.

#Lodging and meals are included in here, which is interesting; both are from John B. Larson.
```

