---
title: "Reverse Engineering Project"
author: "Stella S. Canino, Wade DeVinney, and Miles Grovic"
date: "October 20, 2024"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story in this Washington Post article: "Lawmakers expensed millions in 2023 under new program that doesn’t require receipts."

Link: https://www.washingtonpost.com/politics/2024/06/04/house-representatives-expenses-receipts-lodging/


For this project, we considered that the article reported on the expenses of 2023 and looked for the CSV files in the archives of the U.S. House of Representatives. The entire year of 2023 were divided into four csv files that
we had to join together to form one csv file with the whole year. This data required some cleaning by turning the BUDGET.OBJECT.CODE into a character. Then, we cleaned the dates columns of the CSV file. 

For the sentences we reversed, the engineer most involved the expenses of the congresspeople in total. To make it easier to observe, we created a new data frame by filtering by organization, where the name of their office (so them) is. After filtering for their office, we filtered further by description and budget object code. Description is the specific ID number for the types of expenses that are reported. For example, 2109 is the ID for meals and 2107 is the ID for Lodgings. We know these codes mean what they mean via the description column. We have a mix and match of this depending on the preference of the person answering the sentence.

After that, we would group by the budget object code or description and then summarize by the month amount and total amount in descending order, depending on the question.


## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}

```

```{r}
# Turn off scientific notation
options(scipen=999)

library(tidyverse)
library(janitor)
library(lubridate)

```


## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

april_june <- read.csv("~/Documents/GitHub/data_jour_2024_fall_SSCQ/major_assignments/reverse_engineering/data/APRIL-JUNE 2023 SOD DETAIL GRID-FINAL.csv")

jan_march <- read.csv("~/Documents/GitHub/data_jour_2024_fall_SSCQ/major_assignments/reverse_engineering/data/JAN-MAR-2023-SOD-DETAIL-GRID-FINAL.csv")

july_sept <- read.csv("~/Documents/GitHub/data_jour_2024_fall_SSCQ/major_assignments/reverse_engineering/data/JULY-SEPTEMBER-2023-SOD-DETAIL-GRID-FINAL.csv")

oct_dec <- read.csv("~/Documents/GitHub/data_jour_2024_fall_SSCQ/major_assignments/reverse_engineering/data/OCT-DEC-2023-SOD-DETAIL-GRID-FINAL.csv")

# Clean required data and prepare for analysis if needed. 
```

```{r}

april_june$BUDGET.OBJECT.CODE <- as.character(april_june$BUDGET.OBJECT.CODE)
jan_march$BUDGET.OBJECT.CODE <- as.character(jan_march$BUDGET.OBJECT.CODE)
july_sept$BUDGET.OBJECT.CODE <- as.character(july_sept$BUDGET.OBJECT.CODE)
oct_dec$BUDGET.OBJECT.CODE <- as.character(oct_dec$BUDGET.OBJECT.CODE)

# Now bind the rows together
us_rep_expense <- bind_rows(april_june, jan_march, july_sept, oct_dec)

glimpse(us_rep_expense)
```

```{r}
us_rep_expense <- us_rep_expense |>
  mutate(
    PERFORM.START.DT = as.Date(PERFORM.START.DT, format = "%d-%b-%y"),
    PERFORM.END.DT = as.Date(PERFORM.END.DT, format = "%d-%b-%y"),
    TRANSACTION.DATE = as.Date(TRANSACTION.DATE, format = "%d-%b-%y")
  )
```

```{r}
glimpse(us_rep_expense)
```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: P1. Rep. Jack Bergman (R-Mich.) was the program’s top spender last year. He was reimbursed more than $32,000 for lodging and nearly $12,000 for meals in 2023, according to data released by the House as of Tuesday.

* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

For Rep. Jack Bergman we found that he spent $49,695.49 in Lodgings in 2023, which
matches the description of more than $32,000. However, for meals we were only able
to calculate 4388.89. We believe that for meals may have been reclassified under
something else to get this 12,000 total. 

```{r}
# Put code to reverse engineer sentence here

JACK <- us_rep_expense |> filter(ORGANIZATION == "2023 HON. JACK BERGMAN") 


# Display results of code below this codeblock
#This is Meals
JACK |>
  filter(BUDGET.OBJECT.CODE %in% c("2110", "2603", "2602", "2109")) |>
  group_by(ORGANIZATION, BUDGET.OBJECT.CODE) |>
  summarise(TOTAL_AMOUNT = sum(AMOUNT)) |>
  arrange(desc(TOTAL_AMOUNT))

```


```{r}
##This is lodging for Bergman
JACK |> filter(BUDGET.OBJECT.CODE == "2105") |>
  group_by(ORGANIZATION, BUDGET.OBJECT.CODE) |>
  summarise(TOTAL_AMOUNT = sum(AMOUNT)) |>
  arrange(desc(TOTAL_AMOUNT))

```

### Sentence 2

* **Sentence text**: Rep. Matt Gaetz (R-Fla.) was the program’s second-highest overall spender, according to filings released as of Tuesday, with over $30,000 in lodging expenses and more than $11,000 for meals in 2023. 

* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

For Rep. Matt Gaetz, we found that for lodgings he spent $36,340.09 in 2023,
which fits the description of over $30,000 in the article. For meals, we have
$18411.61, which fits the description of spending more than $11,000 on meals.

```{r}
# Put code to reverse engineer sentence here
##GAETZ LODGINGS

Gaetz <- us_rep_expense |> filter(ORGANIZATION == "2023 HON. MATT GAETZ") 

# Display results of code below this codeblock

##This is for lodgings
Gaetz |>
  filter(BUDGET.OBJECT.CODE == c("2107", "2105")) |>
  group_by(ORGANIZATION, BUDGET.OBJECT.CODE) |>
  summarise(TOTAL_AMOUNT = sum(AMOUNT)) |>
  arrange(desc(TOTAL_AMOUNT))

# Display results of code below this codeblock

```

```{r}
#Meals 
Gaetz |>
  filter(BUDGET.OBJECT.CODE %in% c("2109", "2110", "2603", "2602")) |> 
 group_by(ORGANIZATION, BUDGET.OBJECT.CODE) |>
  summarise(TOTAL_AMOUNT = sum(AMOUNT)) |>
  arrange(desc(TOTAL_AMOUNT))

```


### Sentence 3

* **Sentence text**: Some members of Congress who own homes in the Washington area, including Reps. Patrick T. McHenry (R-N.C.), Ro Khanna (D-Calif.) and Michael McCaul (R-Tex.), have chosen not to participate in the program at all. Rep. Jim Banks (R-Ind.), who owns a $1 million home in Virginia, was reimbursed less than $1,500 each month.

* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

We were able to confirm that Jim Banks was reimbursed less than $1,500 a month of lodgings, 
which was used to compare Gaetz lodgings expense. The highest expense in a month
that he had was 1401.92 in October. 

```{r}
# Put code to reverse engineer sentence here
JIM <- us_rep_expense |> filter(ORGANIZATION == "2023 HON. JIM BANKS")
# Display results of code below this codeblock
```

```{r}
#Lodgings per month of JIM
JIM |> filter(BUDGET.OBJECT.CODE == c("2107", "2105")) |>
  mutate(Month = month(as.Date(PERFORM.END.DT, format = "%d-%b-%y`"))) |>
 group_by(Month, BUDGET.OBJECT.CODE) |>
  summarise(AMOUNT) |>
  arrange(desc(AMOUNT))
    
```


### Sentence 4

* **Sentence text**: Mace, who co-owns a $1,649,000 Capitol Hill townhouse she purchased in 2021 with her then-fiancé, Patrick Bryant, expensed a total of $27,817 in 2023. She expensed over $3,000 for lodging in January, March, and May and over $4,000 for lodging in October.

* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

For Nancy Mace, we found that she reimbursed $33,263.22 for lodgings in 2023, which is
more than the total expressed in the article.For the months of January, 
March and May, we confirmed that she reimbursed over 3,000 those months: 3612.00	(MAY),	3354.00	(MARCH)	
and 3196.00 (JAN). We believe that its possible that our lodging numbers were larger
because of a differiation that the authors did with those numbers. 

```{r}
# Put code to reverse engineer sentence here

MACE <- us_rep_expense |>
  filter(ORGANIZATION == "2023 HON. NANCY MACE")

##LODGING
MACE |> filter(DESCRIPTION == "LODGING") |>
  group_by(DESCRIPTION) |>
  summarise(TOTAL_AMOUNT = sum(AMOUNT)) |>
  arrange(desc(TOTAL_AMOUNT))

# Display results of code below this codeblock
```

```{r}

MACE |> filter(DESCRIPTION == "LODGING") |>
  mutate(Month = month(as.Date(PERFORM.START.DT, format = "%d-%b-%y`"))) |>
 group_by(DESCRIPTION) |>
  summarise(AMOUNT, Month) |>
  arrange(desc(AMOUNT)) 
    
    
    

```


### Sentence 5

* **Sentence text**: Swalwell, who purchased a $1,215,000 home in the Eckington neighborhood of Washington, was reimbursed more than $20,000 for lodging expenses in 2023, according to the data released as of Tuesday. In May 2023, he received $2,838 in lodging reimbursement.

* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

For Swalwell, we found that he spent $28,185.23 on Lodgings in 2023, which match the description
of him being reimbursed more than $20,000. Furthermore, we were able to confirm that he spent
$2838.00 in May with our code. 

```{r}
# Put code to reverse engineer sentence here

SWALWELL <- us_rep_expense |> filter(ORGANIZATION == "2023 HON. ERIC SWALWELL")

##TOTAL LODGINGS
SWALWELL |> filter(BUDGET.OBJECT.CODE == c("2105","2107")) |>
  group_by(BUDGET.OBJECT.CODE) |>
  summarise(TOTAL_AMOUNT = sum(AMOUNT)) |>
  arrange(desc(TOTAL_AMOUNT))

# Display results of code below this codeblock

```

```{r}
##Lodging for May for Swawell. 

SWALWELL |> filter(BUDGET.OBJECT.CODE == c("2105","2107")) |> 
  mutate(Month = month(as.Date(PERFORM.END.DT, format = "%d-%b-%y`"))) |>
 group_by(Month, BUDGET.OBJECT.CODE) |>
  summarise(AMOUNT) |>
  arrange(desc(AMOUNT))
  

```



-30-