---
title: "lab_03"
author: "Derek Willis"
date: "2024-08-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse
library(tidyverse)

# Load janitor
library(janitor)
```

## Load data

```{r}
# Load MD slice of FEC House candidate contribution data
md_contribs <- read_rds("data/lab_03.rds")
```

## Answer questions

Q1: Which seven cities in Maryland had the largest *number* of contributions from donors? How many did each have? To answer this question, you'll have to clean the messy "contributor_city" field.

Note: To answer this question, you'll need to see every variation of the "contributor_city" for the top seven cities, which you should identify through some initial analysis. Then find variations of them and change those variations into the most common version.

A1:

The seven cities with largest number of is Bethesda (1177), CHEVY CHASE	(705), BALTIMORE	(692), ANNAPOLIS	(492), POTOMAC	(491),	SILVER SPRING	(408)	and ROCKVILLE	(359).		


```{r}
colnames(md_contribs)
glimpse(md_contribs)

```

```{r}
md_contribs |> 
  group_by(contributor_city) |> 
  summarise(count = n()) |> 
  arrange(desc(count))
          

```

```{r}

 md_contribs <- md_contribs |> 
  clean_names() |> 
  mutate(contributor_city = case_when(
    contributor_city == "BELAIR" ~ "BEL AIR",
    contributor_city == "ANAPOLIS" ~ "ANNAPOLIS", 
    contributor_city == "BLATIMORE" ~ "BALTIMORE",
    contributor_city == "BALRIMORE" ~ "BALTIMORE",
    contributor_city == "4615 NORTH PARK AVE APT. 1410" ~ "CHEVY CHASE",
    contributor_city == "ROCKWELL" ~ "ROCKVILLE",  # Fixed here
    contributor_city == "GAITHERBURG" ~ "GAITHERSBURG",
    contributor_city == "SILVER SPRINGS"  ~ "SILVER SPRING",
    contributor_city == "POTAMIC" ~ "POTOMAC",
    contributor_city == "COLOMBIA" ~ "COLUMBIA",
    TRUE ~ contributor_city))
    
```


```{r}
md_contribs |> 
  group_by(contributor_city) |> 
  summarise(count = n()) |> 
  arrange(desc(count))
            
```


   
Q2: Find the 5-digit ZIP code in Maryland that had the highest total amount of contributions (by dollar value). You will need to use mutate to ensure that you have a zip code column with the first five characters from the existing zip code column.

Where is that zip code and what kind of area is it, demographically? Which committee is the largest recipient of money from that zip code? How many of the top five recipients are Maryland candidates?

A2: 

#WRONG --> 20815. It's looking for amount of people that contributed not the amount they donated. It's Chevy Chase Its Raskin the committee given to. I think it would be 3.

The Zip code with the highest total amount of contributions is 20817. This is the zip code for Montgomery County, which is a very affluent area with a median household income of six figures, and is close to Washington, D.C. The largest recipient of is Helmer for Congress that's Virginia's 10th district. There are only two candidates out of the five of this zip code that are Md. candidates.

```{r}

###This code cleans the zips. 

md_contribs <- md_contribs |>
  clean_names() |> 
  distinct() |> 
  mutate(contributor_zip = as.character(contributor_zip)) |>
  mutate(contributor_zip = str_sub(contributor_zip, start = 1L, end = 5L))


```

```{r}
 #Answer 1 of zip code with most amount $ contributors.
md_contribs |>
  group_by(contributor_zip) |>
  summarise(total_amount = sum(contribution_receipt_amount)) |> 
  arrange(desc(total_amount)) 
```

```{r}

##Which committee is the largest recipient of money from that zip code? 

md_contribs |>
  filter(contributor_zip == "20817") |>
  group_by(committee_name) |>
  summarise(total_amount = sum(contribution_receipt_amount)) |> 
  arrange(desc(total_amount))

```


  Q3: Let's look at top individual donors. Write code to generate a list of top donors using name and city, adding up the total amount of contributions. What does the original data say about the top two donors - who are they and what do they do? Do you see anything in the list of top 10 donors that would make you uncertain about the accuracy of these results?

A3:

The top two donors are Melanie Jacobsen and Mark P. Jacobsen, who are from Bethesda. Mark identifies with three different occupations that makes me question why he's registering himself that way. Another worry is that R put Mark Jacobsen and Drezner twice, skewing the results that make Mark first place as the largest donor by more than a million dollars in contributions.
 

For that reason, I'm unsure of the accuracy of these results. 

```{r}

md_contribs |>
  group_by(contributor_name, contributor_city) |>
  summarise(total_amount = sum(contribution_receipt_amount)) |>
  arrange(desc(total_amount))
```


```{r}

md_contribs |>
  filter(contributor_name == "JACOBSEN, MELANIE"| contributor_name == "JACOBSEN, MARK P."|  contributor_name == "DREZNER, JEFFREY L." |  contributor_name == "KRACOV, DANIEL") |>
group_by(contributor_name, contributor_occupation) |>
  summarise(n=n())

```



 