---
title: "Data Analysis Project"
author: "Stella S. Canino, Wade DeVinney, and Bao Bao"
date: "October 20, 2024"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
# Turn off scientific notation
options(scipen=999)
#Turn on libraries
library(tidyverse)
library(janitor)
library(lubridate)

```

*Cleaning and Preparation of Data*

```{r}
#This is the csv files of all the disbursements published in 2024.

april_june24 <- read.csv("data/APRIL-JUNE-2024-SOD-DETAIL-GRID-FINAL.csv")

jan_march24 <- read.csv("data/JAN-MAR-2024-SOD-DETAIL-GRID-FINAL.csv")

# Clean required data and made turned this into a character or it wouldn't run.
april_june24$BUDGET.OBJECT.CODE <- as.character(april_june24$BUDGET.OBJECT.CODE)
jan_march24$BUDGET.OBJECT.CODE <- as.character(jan_march24$BUDGET.OBJECT.CODE)

# Now bind the rows together
us_rep_expense24 <- bind_rows(april_june24, jan_march24)

glimpse(us_rep_expense24)

```

```{r}
#Now, the next thing is to clean this further by format the dates from integers.
us_rep_expense24 <- us_rep_expense24 |>
  mutate(
    PERFORM.START.DT = as.Date(PERFORM.START.DT, format = "%d-%b-%y"),
    PERFORM.END.DT = as.Date(PERFORM.END.DT, format = "%d-%b-%y"),
    TRANSACTION.DATE = as.Date(TRANSACTION.DATE, format = "%d-%b-%y")
  )

glimpse(us_rep_expense24)

```

```{r}
#Now, we further this dataframe further to show only the congressmen 

US_congresspeeps_24 <- us_rep_expense24 |> filter(FISCAL.YEAR.OR.LEGISLATIVE.YEAR == "LY2024") |> 
    filter(str_detect(ORGANIZATION, "HON")) |> filter(!str_detect(DESCRIPTION,"TOTALS"))  

glimpse(US_congresspeeps_24)
```

**Data Analysis Section**

Q1: Who is the top spender based on the data available for 2024, and what have they spent the most on?

As of July 2024, the top congressmen that used the disbursement program is 
Alexander X. Mooney of West Virginia's 3rd district. Mooney spent a total of
$1,276,864.48. He spent the most in advertising with $276,950.18. 

We did a simple group_by and summarize by the total amount. To get an accurate
total amount we first filtered out by "TOTALS", which skewed our results on the
top of this code. Then we created a new dataframe to only observe Mooney. In the
next question, we analyzed vendor names and had it initially show one billion with
NAs. With further inspection, we saw that there was a Totals section and other 
exchanges that increased the results.

As a result we filtered by string detect (Outside) and (Transfer) as well.

I believe it's newsworthy that he spent more than a quarter of his disbursements
on advertising, since this is an election year. For context, Mooney was running
for the Republican primary of West Virginia to fill the vacancy of Sen. Joe
Manchin, and he lost the primary to Gov. Justice. 

#It is illegal to use disbursements for the reelection campaigns of congresspeople.
 What if the advertisements are notes from the office on all the victories and 
bills he formed part of as a way to inform his constituents? Framed in that way,
a representative could indirectly use this program as a way to aid his campaign
for the primary race.

```{r}
#1A
US_congresspeeps_24 |> 
  group_by(ORGANIZATION) |>
  summarise(Total_Amount = sum(AMOUNT, na.rm = TRUE)) |>
  arrange(desc(Total_Amount)) 
```

```{r}
#1b
Mooney24 <- US_congresspeeps_24 |>
  filter(ORGANIZATION == "2024 HON. ALEXANDER X. MOONEY")

Mooney24 |>
group_by(DESCRIPTION) |>
summarise(Total_Amount = sum(AMOUNT, na.rm = TRUE)) |>
  arrange(desc(Total_Amount)) |> filter(!str_detect(DESCRIPTION,"TOTALS")) |>
  filter(!str_detect(DESCRIPTION,"(TRANSFER)")) |>
   filter(!str_detect(DESCRIPTION,"(OUTSIDE)"))
```

Q2: Who is the top recipient of government money?

As explained above, we had a lot of NAs when investigating the top recipient who is
paid by congressmen. We made to separate dataframes to investigate this further. 
Since we separated this, we were able to investigate the named recipients easier.

We found that the top recipient is the U.S. Postal Service(1) and Citibank(2),
LEIDOS DIGITAL SOLUTIONS INC(4) which isn't very interesting.
However, the interesting top vendors are Amplify Inc(3) and AEJ Group LLC (5), and
The Franking Group (7). These three companies are related to media relations and
advertising firms that are working with congresspeople

This is newsworthy since we see this pattern from Q1 advertising-related
expenses are the most that are being reimbursed in this legislative year. This 
is important because it could highlight how congresspeople are using 
their media and advertising disbursements to reach voters and change public opinion
before their primary or general election.

```{r}
# Separate rows with missing or blank vendor names
missing_vendors <- US_congresspeeps_24 |> 
  filter(is.na(VENDOR.NAME) | str_trim(VENDOR.NAME) == "")

# Examine rows with valid vendor names
valid_vendors <- US_congresspeeps_24 |> 
  filter(!is.na(VENDOR.NAME) & str_trim(VENDOR.NAME) != "")

#There are more NA vendors that named vendors. --> with valid vendors having 136,086 obs.
```

```{r}
valid_vendors |> 
  group_by(VENDOR.NAME) |>
  summarise(Total_Amount = sum(AMOUNT, na.rm = TRUE)) |>
    arrange(desc(Total_Amount))
```

Q3: Who is spending the most money on advertising?

The congressmen who spent the most money on advertising is Hon. Shri Thanedar
of Michigan's 13th district with $584,942.46. 

It's newsworthy to Thanedar had one of the most competitive primaries in his state
going against two nominees, and he won it by 54%. We could also try to observe
how has the advertising expense of Thanedar changed compared to last year 
and during the months leading up to the primary(August), and general election (November).

Another thing worth investigating is why Thanedar also appeared as a vendor
that's being given money as well. 

```{r}
US_congresspeeps_24 |>
  filter(DESCRIPTION == "ADVERTISEMENTS") |> 
  group_by(ORGANIZATION) |>                
  summarise(Total_Amount = sum(AMOUNT, na.rm = TRUE)) |> 
  arrange(desc(Total_Amount))   
```

Q4:Who is the top spender for traveling (airplanes, cars, etc) of 2024 available data?

The top spender for traveling disbursements is Hon. Lance Gooden of Texas's
5th district with $143,418.52. As this is an election year, such high spending
is noteworthy if that amount of travel is related to campaigning for himself or 
others for the primary or general election. 

```{r}
#Q2 Who is the top spender for traveling (airplanes, cars, etc)  in what is of  2024?

filtered_data <- US_congresspeeps_24 |> 
  filter(
    SORT.SUBTOTAL.DESCRIPTION == "TRAVEL") 
  
filtered_data <- filtered_data |> 
  filter(!DESCRIPTION %in% c("MEALS", "LODGING", "TRAVEL TOTALS:", "WI-FI ON TRAVEL"))

filtered_data |>
  group_by(ORGANIZATION) |>
 summarise(Total_Amount = sum(AMOUNT, na.rm = TRUE)) |> 
  arrange(desc(Total_Amount))   


```

 Q5 Which congressmen spent the most on advertising on the data available of 2024 by month?

The congressmen who spent the most on advertising by month is Shri Thanedar in
July with a total of $440,728.26. It's important to note that the primary he 
was facing was in August 5th, which makes this amount in June the majority
of his total advertising expense that's reported. We would have to see what
this increase holds when the July-September data roles around. 

Another important thing to note that the second person that spent the most on
advertising in 2024 was Claudia Tenney of New York who faced a very contentious
primary and comes from a district that has flipped a lot. Her largest expense was
in May 2025 with $210,102. The primary race was in June, meaning that this 
expense was one month before her election.

This is newsworthy since we see a pattern of the advertising expenses
increasing near the months of the congressmen primaries. A further investigation
could reveal more patterns of this use of reimbursements to help camgaigns.

```{r}
US_congresspeeps_24 |>
  filter(DESCRIPTION == "ADVERTISEMENTS") |> 
  mutate(Month = format(as.Date(TRANSACTION.DATE), "%Y-%m")) |> # Extract year-month from the date
  group_by(ORGANIZATION, Month) |>                
  summarise(Total_Amount = sum(AMOUNT, na.rm = TRUE)) |> 
  arrange(desc(Total_Amount))

```



