---
title: "lab_04"
author: "Stella S. Canino"
date: "2024-09-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About this lab

To complete this lab, you need to: \* run existing code as directed (look for **Task**). \* modify existing code as directed (look for **Task**). \* write code in empty codeblocks provided to answer questions included (look for **Q**). \* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings

You'll need to load three packages for this: the tidyverse, lubridate and janitor.

**Task** load these three packages.

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
library(janitor)
library(lubridate)

```

For this lab, we want to investigate spending by the two leading Senate candidates in Maryland during this election, Angela Alsobrooks and Larry Hogan. Our goal is to try and standardize the purpose of the expenditures so that we can compare the two campaigns. We have the data, but it's messy - purpose descriptions are spelled differently and capitalized differently - and we need to clean it up. We'll use a combination of RStudio and OpenRefine to do that.

The basic process here is to start in RStudio, export data so that we can use OpenRefine to clean it up, and then bring it back into RStudio to finish our analysis.

## Load Data

You'll need to load one data set, a CSV file of campaign expenditures located in the data folder called "md_senate_expenses.csv"

**Task** Create a codeblock below, then read the data in and assign it to an appropriate variable name. You'll want to make sure that any date columns are actually date datatypes.

```{r}
# Load the Maryland expenditures data table

md_senate_expenses <- read_csv("data/md_senate_expenses.csv")
glimpse(md_senate_expenses)

```

## Answer questions

**Q1.** You've been assigned to compare the spending priorities of the two campaigns, but they describe their spending differently using the `purpose` column. Using OpenRefine, create a project using the original CSV file, make a copy of the `purpose` column called `purpose_clean` and then standardize it, focusing on the purposes that appear most often or have the largest total amounts. You can transform the data as you want - making all records the same case, for example, to help. The goal is to be able to group together as many similar purpose values, so you can choose to call all transportation-related expenses the same thing.

How do you know when you're done? It's a judgment call, but you should be close to 100 unique values for `purpose_clean`, and maybe even below that number.

Then, when you are done standardizing `purpose_clean`, export it as a new CSV file to your data folder, *giving it a different name* and read it back into RStudio as a new dataframe.

Then, using that `purpose_clean` column, write code to find the largest amounts of spending on categories by both campaigns.

**A1.**

The largest amount of spending by both campaigns is Media advertising with 6,438,339.18. 


```{r}
md_sen_new <- read.csv("data/md_sen_reimbursement2.csv")
head(md_sen_new)

```


```{r}

md_sen_new <- md_sen_new |>
  mutate(purpose_clean = case_when(
    purpose_clean == "AD BUY" ~ "Media Advertising",
    purpose_clean == "Online Advertising" ~ "Media Advertising",
       purpose_clean == "PRINT ADS" ~ "Media Advertising",
    purpose_clean == "ADS" ~ "Media Advertising",
        TRUE ~ purpose_clean))


```


```{r}

md_sen_new |>
  group_by(purpose_clean) |>
  summarise(
    count = n(),
    total_amount = sum(amount)
  ) |>
  arrange(desc(total_amount)) 

```

**Q2.** Let's do some research on the top purposes from A1. Feel free to search the Web to understand those terms better if need be, and then write a description comparing and contrasting the two campaigns' largest categories of spending.

**A2.**
Both campaigns share that their top two largest expense is Media Advertising and paying their staff. Alsobrooks has paid two million more for media advertising than Hogan, which is interesting. In contrast, the Alsobrooks spent less in fundraising consulting than Hogan, which was his third largest expense. Another thing is that Alsobrooks spent more money on digital consulting than Hogan, but he spent more money on strategy consulting than she did.

```{r}
#Hogan's money spending $$$

md_sen_new |>
  filter(candidate=="Hogan") |>
  group_by(purpose_clean) |>
  summarise(
    count = n(),
    total_amount = sum(amount)
  ) |>
  arrange(desc(total_amount)) 

```


```{r}
md_sen_new |>
  filter(candidate=="Alsobrooks") |>
  group_by(purpose_clean) |>
  summarise(
    count = n(),
    total_amount = sum(amount)
  ) |>
  arrange(desc(total_amount)) 

```


**Q3.** Pick one of the top categories for each campaign and write code that shows how that spending has changed over time, if at all. You'll need to use lubridate for this. Write a couple of sentences describing that pattern.

For the Alsobrooks campaign, the month with the largest amount of expenses was April (2,651,938) and May (1,535,542) 2024, which were the months leading up to her primary election. In general, the majority of her big expenses was in 2024. For Hogan, the majority of his expenses were also in April and May, which were at the time of his primary. The interesting thing is that the expenses recorded for Hogan started in Feb. 2024, while Alsobrooks dates to May of 2023. Alsobrooks expenses are then more spread out than Hogan.

**A3.**

```{r}

md_sen_new |> 
  filter(candidate == "Alsobrooks") |>
  mutate(date = ymd(date),
   Month = floor_date(date, "month")) |>  
  group_by(Month) |>
  summarise(
    count = n(),
    total_amount = sum(amount)
  ) |>
  arrange(desc(total_amount))
```

```{r}
md_sen_new |> 
  filter(candidate == "Hogan") |>
  mutate(date = ymd(date),
   Month = floor_date(date, "month")) |>  
  group_by(Month) |>
  summarise(
    count = n(),
    total_amount = sum(amount)
  ) |>
  arrange(desc(total_amount))
```


**Q4.** Which committee has the larger payroll or salary spending so far? How many unique recipients of payroll/staff spending are there for each candidate?

For Alsobrooks campaign, the committee with the largest payroll is Gustos for 43,6720, who is the only committee. However, her campaign pays salary to 319 employees for a total of 822015. However, the Hogan campaign has no committee that payrolls or specific individuals that it gives salary or payroll too.

**A4.**

```{r}
md_sen_new |> 
  filter(purpose_clean=="Salary") |>
  filter(candidate =="Hogan") |>
  group_by(organization_name) |>
  summarise(
    count = n(),
   total_amount = sum(amount)
  ) |>
  arrange(desc(total_amount))

md_sen_new |> 
  filter(purpose_clean=="Payroll Taxes") |>
  filter(candidate =="Hogan") |>
  group_by(organization_name) |>
  summarise(
    count = n(),
   total_amount = sum(amount)
  ) |>
  arrange(desc(total_amount))


```

```{r}
##This is for the committee with payroll taxes.

md_sen_new |> 
  filter(purpose_clean=="Payroll Taxes") |>
  filter(candidate =="Alsobrooks") |>
  group_by(organization_name) |>
  summarise(
    count = n(),
   total_amount = sum(amount)
  ) |>
  arrange(desc(total_amount))

### This is for salary
md_sen_new |> 
  filter(purpose_clean== "Salary") |>
  filter(candidate =="Alsobrooks") |>
  group_by(organization_name) |>
  summarise(
    count = n(),
   total_amount = sum(amount)
  ) |>
  arrange(desc(total_amount))



```


**Q5** What's the most interesting story idea you've found in this data?

The most interesting story idea I found was why did Hogan start reporting expenses in February and what's the strategy intended behind it.

**A5**
